---
title: "3. QC analysis via peaks detection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3. QC analysis via peaks detection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(crayon.enabled=F)
```

# Peak analysis

`CNAqc` can perform quality control (QC) of somatic mutations, allele-specific CNAs and tumour purity. 3 distinct procedures are available, all called by function `analyze_peaks`. The idea is as follows: a CNA segment with $m$/ $M$ copies of the minor/ major alleles is consistent with mutation data if VAFs peak at certain positions we can compute. CNAqc computes expected peaks, and compares them to the one it detects from data. The theory works for both clonal and subclonal segments, and a purity error parameter (`purity_error`) determines the maximum error we can tolerate before we fail a sample.

```{r setup}
library(CNAqc)

# We work with the template dataset. 
data('example_dataset_CNAqc', package = 'CNAqc')

x = CNAqc::init(
  example_dataset_CNAqc$snvs, 
  example_dataset_CNAqc$cna,
  example_dataset_CNAqc$purity,
  ref = 'hg19')

x = CNAqc::example_PCAWG

print(x)
```


We run it with default parameters, otherwise one can select only certain segments  to fit or change purity error tolerance etc.
```{r, fig.width=11, fig.height=3, warning=F}
x = analyze_peaks(x)
```
An S3 methods print to screen all QC results based on peak-detection.

```{r, fig.width=11, fig.height=3}
print(x)
```


## Simple clonal CNAs

This QC is the most important as it can be used to detect adjustments required to better fit purity or ploidy of the sample (to re-run the copy number caller), or to select among alternative copy number solutions returned from a caller (tetraploid versus triploid).

For these segments the package computes a linear score from the distance between data peaks, and their theoretical expectation. The peaks are determined via:

- a kernel-density (unitary adjustment, Gaussian kernel) used to smooth the data distribution, and the   [peakPick](https://cran.rstudio.com/web/packages/peakPick/index.html) package to determine peaks in the smoothed density.

- a Binomial mixture model using [Bmix](https://caravagnalab.github.io/BMix/).

There are two possible algorithmic strategies to match peaks with clonal simple CNAs (configured by parameter `matching_strategy`): one matches the closest peaks as defined by euclidean distance, the other ranks peaks  right to left, and prioritises right-most peaks (see the manual for more details). A call to `plot_peaks_analysis` with default parameters shows the results from peak detection.

```{r, fig.width=11, fig.height=3, warning=F}
plot_peaks_analysis(x)
```

Plots are ordered by fixed positions (gray panels represent karyotypes that have not been used). Each vertical dashed line is the expected peak, the bandwidth around being the tolerance we use to match peaks (based on `purity_error`, adjusted for segment ploidy and tumour purity). Each dot is a peak detected from data, with a bandwidth of tolerance (fixed) around it. The colours green/red reflect if a peak is matched or not; the overall QC status for a karyotype  is given by the colour of the facet and the box surrounding the whole figure assembly reports the QC status for the sample.

Plots can also be generated without assembly or empty plots (see the parameters of function `plot_peaks_analysis`).

## Complex clonal CNAs

The QC procedure for these CNAs uses only the kernel density estimation to make the computation faster.
```{r, fig.width=8, fig.height=5, warning=F}
plot_peaks_analysis(x, what = 'general')
```
The plot reports dashed lines for expected peaks (for mutation with multiplicity from 1 to the maximum of the Major/minor allele counts), and dots for each data peak.The colours green/red report the QC status for each peak; with these CNAs no karyotype-level or sample-level scores are produced.

##  Subclonal CNAs 

The QC procedure for these CNAs uses only the kernel density estimation (as for complex CNAs), and considers 2 subclones with distinct mixing proportions. Differently from clonal CNAs, however, here the analysis is carried out at the level of each segment (without pooling segments with the same karyotypes) so to support callers that report segment-specific CCF values (e.g., Battenberg).

```{r, fig.width=8, fig.height=12, warning=F}
plot_peaks_analysis(x, what = 'subclonal')
```
The visual layout of this plot is the same of complex clonal CNAs; not that the facet reports the distinct evolutionary models that have been generated to QC subclonal CNAs.


# Summary results

For every type of copy number analyzed several information are reported, tables with summary peaks are available inside `x$peaks_analysis`. More precisely
```{r, fig.width=3, fig.height=3}
# Simple clonal CNAs
x$peaks_analysis$matches

# Complex clonal CNAs
x$peaks_analysis$general$expected_peaks

# Subclonal CNAs
x$peaks_analysis$subclonal$expected_peaks
```

The most helpful table is usually the one for simple clonal CNAs `x$peaks_analysis$matches`, which reports several information:

- `mutation_multiplicity` and `karyotype`, which are the number of copies of the mutation, and the segment's karyotype. Each segment with `discarded = FALSE` has been analysed;
- `peak`, `x`, `y` report where the expected peak should be found, and where it has been found (`x` and `y`);
- `offset`, `weight` and `score` that report the factors of the final `score`;
- `QC` that reports if the  peak is considered matched.

The overall sample-level QC result - `"PASS"/"FAIL"` - is available in
```{r}
x$peaks_analysis$QC
```

You can visually summarise QC results for all the analyses available in CNAqc (peaks-based and [CCF-based](https://caravagn.github.io/CNAqc/articles/ccf_computation.html)).

```{r, fig.width=6, fig.height=10, message=F, warning=F}
plot_qc(x)
```


# Per-chromosome analysis

CNAqc analyses can be run per chromosome. First, split your data by chromosome using function `split_by_chromosome`, and then run the analysis for every chromosome. 

