---
title: "5. Smoothing copy number segments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5. Smoothing copy number segments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(crayon.enabled=FALSE)
```

```{r setup}
library(CNAqc)
```

Smoothing merges, for each chromosome, all contiguous segments that have the same minnor and major allele counts, and are separated by up to $\delta$ base pairs (default $\delta = 10^6$, 1 megabase). 

Only clonal CNAs can be smoothed; subclonal CNAs are retained  but cannot be smoothed because the CCF of contiguous segments might be different.

```{r echo=FALSE, results=TRUE, message=FALSE}
# Dataset available with the package 
data('example_dataset_CNAqc', package = 'CNAqc')

x = CNAqc::init(
  mutations = example_dataset_CNAqc$mutations, 
  cna = example_dataset_CNAqc$cna,
  purity = example_dataset_CNAqc$purity,
  ref = 'hg19')
```



```{r, fig.width=11, fig.height=3}
# Before smoothing
print(x)

# After smoothing
x = smooth_segments(x, maximum_distance = 1e6) # default
```

The old CNAqc object is retained inside the new one
```{r, fig.width=11, fig.height=3}
print(x$before_smoothing)

# The new one should have fewer segments
print(x)
```

You can visualise the effect of the smoothing.
```{r, fig.width=15, fig.height=7}
plot_smoothing(x)
```

```{r, fig.width=6, fig.height=6}
plot_multisample_CNA(list(`Before` = x$before_smoothing, `After` = x))
```

