---
title: "5. Smoothing copy number segments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5. Smoothing copy number segments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(crayon.enabled=FALSE)
```

```{r setup}
library(CNAqc)
```

We work with the template dataset. 

```{r echo=FALSE, results=TRUE, message=FALSE}
# Dataset available with the package 
data('example_dataset_CNAqc', package = 'CNAqc')
x = CNAqc::init(
  example_dataset_CNAqc$snvs, 
  example_dataset_CNAqc$cna,
  example_dataset_CNAqc$purity,
  ref = 'hg19')

print(x)
```


Smoothing works with a simple algorithm that merges, inside each chromosome, all contiguous segments that have the same absolute copy number values (minnor and major allele counts), and are separated by up to $\delta$ base pairs. 

The default value of $\delta = 10^6$ (1 megabase).

```{r, fig.width=11, fig.height=3}
# Before smoothing
print(x)

# After smoothing
x = smooth_segments(x, maximum_distance = 1e6) # default
```

The old obbject is retained inside the new one
```{r, fig.width=11, fig.height=3}
print(x$before_smoothing)

# The new has theoretically smaller segments
print(x)
```

You can visualise the effect of the smoothing automatically using `CNAqc` functions.
```{r, fig.width=15, fig.height=7}
plot_smoothing(x)
```

```{r, fig.width=6, fig.height=6}
plot_multisample_CNA(list(`Before` = x$before_smoothing, `After` = x))
```

