---
title: "Genome fragmentation analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Genome fragmentation analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(crayon.enabled=FALSE)
```

```{r setup}
library(CNAqc)
```

We work with the [template dataset](https://caravagn.github.io/CNAqc/articles/Introduction.html). 


```{r echo=TRUE, results=FALSE, include=FALSE}
data('example_dataset_CNAqc', package = 'CNAqc')
x = init(example_dataset_CNAqc$snvs, example_dataset_CNAqc$cna,example_dataset_CNAqc$purity)
```

### Fragmentation of individual arms

The fragmentation of a chromosome arm is assessed with a statistical test based on counting
the size of the copy number segments mapping to the arm. 

```{r, fig.width=3, fig.height=3}
# A histogram of segments' lenght
plot_segment_size_distribution(x)
```
 
`CNAqc` counts, for every arm with lenght $L$ nucleotides: 
 
 * $n_s$, the number of mapped CNA segments shorter than $\delta\%$ of $L$;
 * $n_l$, the number of mapped CNA segments longer than $\delta\%$ sof $L$.

A one-sided Binomial test is used to compute a p-value for the null hypothesis of seeing $n_s$ observations in $n_l$ trials, assuming a Binomial success probability $p = \delta > 0$.  $p$ represents a model where each segment length is equally likely (uniform distribution). 

In this way the test accounts for the difference in lenghts of the chromsome arms; a p-value per arm is reported and adjusted for multiple hyoptheses (Bonferroni). 

```{r, fig.width=3, fig.height=3}
# Test with default parameters (small segments: < 20% of chromosome arm)
x = detect_arm_overfragmentation(x)

print(x)
```

You can produce a arm-level report for the fragmentation test, with: 

* a scatter of the counts per arm, with scaled the p-values;
* a jump statistics per arm, $J$. 

$J$ is the sum of the variation in total copy number profiles, evaluated among each pair of contiguous segments.

Significantly overfragmented arms with high $J$  have a "scattered" copy number profile. Those with low $J$ are more uniform, as they show little no copy number change, and can be possibly smoothed (see below).

```{r, fig.width=7, fig.height=12}
plot_arm_fragmentation(x, zoom = 3)
```

Once available, these results appear in any call to `plot_segments` as annotated purple squares sorrounding the arms.

```{r, fig.width=9.5, fig.height=3}
# Default plot has now segments
plot_segments(x)
```

[Smoothing](https://caravagn.github.io/CNAqc/articles/smoothing.html) is a good way to start cleaning up the fragmented sets of arms, as it merges segments that have no karyotype change, and are closer than a certain distance.

```{r, fig.width=9.5, fig.height=3}
# Smooth with default parameters
x = smooth_segments(x)

# Re-compute the fragmentation
x = detect_arm_overfragmentation(x)

print(x)
```

```{r, fig.width=10, fig.height=12}
plot_arm_fragmentation(x, zoom = 6)
```





