---
title: "QC analysis via peaks detection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{QC analysis via peaks detection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(crayon.enabled=F)
```



`CNAqc` can determine if purity and CNA segments "fit well" to mutation data, an hortogonal quality check (QC) metric that can be used to adjust purity/ ploidy.

 `CNAqc`  matches allelic imbalance of CNA segments to the allelic frequencies of somatic mutations. The package implements a karyotype-weighted linear score which uses the distance between data peaks (empirical), and their theoretical expectation. This score accouts for normal plodiy (germline), and tumour purity/  ploidy. The peaks are determined via kernel density estimation and peak-detection heuristics.

```{r setup}
library(CNAqc)
```

We work with the [template dataset](https://caravagn.github.io/CNAqc/articles/Introduction.html). 

```{r echo=FALSE, results=FALSE, message=FALSE, include=FALSE}
# Dataset available with the package -- see Getting Started.
data('example_dataset_CNAqc', package = 'CNAqc')
x = CNAqc::init(
  example_dataset_CNAqc$snvs, 
  example_dataset_CNAqc$cna,
  example_dataset_CNAqc$purity)
```

Peak detection routines can are implemented in function `analyze_peaks`. 

The idea is as follows: a segment with $m$ and $M$ copies of the minor and major alleles is consistent to mutation data if the  VAF distribution peaks at certain values, which we can compute. Profiling then of the VAF distribution with a peak-detection heuristic allows `CNAqc` to measure the if peaks match. Large deviance suggests usually wrong purity estimation.

# Peak analysis

```{r, fig.width=11, fig.height=3, warning=F}
# Run default analysis
x = analyze_peaks(x)
```

`CNAqc` analyze  LOH regions (A, AA), diploid regions (AB), and amplification regions (AAB, AABB). These correspond to
`'1:1'`, `'2:1'`, `'2:0'` and `'2:2'` in "Major:minor" notation. Karyotypes are subset by their size (normalized for
the number of input mutations), and by default karyotypes smaller than 5% of the actual mutational burden are not considered

A Kernel-density method (unitary adjustment, Gaussian kernel) is used to smooth the data distribution, and the  [peakPick: Peak Picking Methods Inspired by Biological Data](https://cran.rstudio.com/web/packages/peakPick/index.html) package is used   to determine peaks in the smoothed density.  Peaks below a mnimum 
density (noise and miscalled segments) are filtered, and the others are matched to a predefined location with some tolerance (e.g., plus/minus `0.5 * epsilon`, where `epsilon = 0.015` is the default).

Theis analysis produces a set of scores that account for karyotype size and peaks offset, and that can be combined into a linear model to compute an overall QC metrics, called `rho`. 
```{r, fig.width=11, fig.height=3}
print(x)
```

A summary table is stored inside the named list `peaks_analysis`.

# Plotting results 

For every karyotype, a plot with the estimated KDE and the matched peaks is available (assembled `ggpubr` figure with); gray panels represent karyotypes that have not been used.

```{r, fig.width=11, fig.height=3, warning=F}
plot_peaks_analysis(x)
```

Individual karyotypes plots are available inside the named lists `peaks_analysis$plots`, named afte each required
karyotype.

```{r, fig.width=3, fig.height=3}
# Tetraploid segments
x$peaks_analysis$plots$`2:2`
```

You can integrate plots from QC analysis with data.
```{r, fig.width=11, fig.height=7, warning=F}
cowplot::plot_grid(
  plot_peaks_analysis(x),
  plot_segments(x),
  rel_heights = c(.8, 1),
  nrow = 2
)
```

# Summary results and Quality Check (QC)

The summary technical table is available as `peaks_analysis$matches` inside `x`. 

```{r}
x$peaks_analysis$matches
```

The table reports several information. The following might be of some use after the analysis:

- `mutation_multiplicity` and `karyotype`, which are the number of copies of the mutation, and the segment's karyotype. Each segment with `discarded = FALSE` has been analysed;
- `peak`, `x`, `y` report where the expected peak should be found, and where it has been found (`x` and `y`);
- `offset`, `weight` and `score` that report the factors of the final `score`;
- `QC` that reports if the  peak is considered matched.

The overall `QC` result - `"PASS"/"FAIL"` - is available in `x`.
```{r}
x$peaks_analysis$QC
```


You can visually summarise the result of the QC analysis, for both peaks and [CCF computation](https://caravagn.github.io/CNAqc/articles/ccf_computation.html).

```{r, fig.width=3, fig.height=3, message=F, warning=F}
plot_qc(x)
```

Here since CCF have not been computed, the only available QC tests are for peak-detection.

