---
title: "Smoothing copy number segments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Smoothing copy number segments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(crayon.enabled=FALSE)
```

```{r setup}
library(CNAqc)
```

We work with the same object shown in the [introductory vignette](file:///Users/gcaravagna/Documents/GitHub/CNAqc/docs/articles/CNAqc.html). 

```{r echo=TRUE, results=FALSE}
# Dataset available with the package -- see Get Started.
data('example_dataset_CNAqc', package = 'CNAqc')
x = CNAqc::init(example_dataset_CNAqc$snvs, example_dataset_CNAqc$cna,example_dataset_CNAqc$purity)
```

Smoothing works with a simple algorithm that merges, inside each chromosome, all contiguous segments that have the same absolute copy number values (minnor and major allele counts), and are separated by up to $\delta$ base pairs. 

The default value of $\delta = 10^6$ (1 megabase).

```{r, fig.width=11, fig.height=3}
# Before smoothing
print(x)

# After smoothing
x = smooth_segments(x, maximum_distance = 1e6) # default
```

The old obbject is retained inside the new one
```{r, fig.width=11, fig.height=3}
print(x$before_smoothing)

# The new has theoretically smaller segments
print(x)
```

You can visualise the effect of the smoothing automatically using `CNAqc` functions.
```{r, fig.width=15, fig.height=7}
plot_smoothing(x)
```


